### 2011-04-15 Text/paper.tex
Almost added the relative velocity picture. Recalculated it for all simulations. Need to add appendix and conclusion.

In comparison with two-dimensional geometries, the experimental works
\cite{shikazono-square,cerro-bubble-train} supported by numerical simulations \cite{heil-threedim,
wang-non-circular} show interesting phenomena in three-dimensional geometries. It was found
\cite{heil-threedim,wong-films} that for rectangular
or square shaped capillaries there is a transition for a
certain capillary number, where the bubble crosssection changes from the non axisymmetric to axis
symmetric case. Non axisymmetric case is attributed to the case where the axial radius is
different from the
diagonal radius and the bubble has the non-circular shape in the channel crossection. In this case
the bubble shape mimics the shape of the square and looks like a rounded square. The dependence
of the diagonal and axial radii on the capillary number is shown in Fig. 
\ref{fig:heil:three:dim}.
\begin{figure}[h]
\includegraphics[width=\textwidth]{Figures/capillary_width_heil.eps}
\caption{\citet{heil-threedim} results for the variation of the deposited liquid for the range of
capillary number. One can see the asymmetry between diagonal and axis diameter for the capillary
number $Ca\leq\hat{Ca}=0.04$. Courtesy of \citet{heil-threedim}. \label{fig:heil:three:dim}}
\end{figure}
The transition number between non-axisymmetric case and symmetric is indicated in a number of
works as $\widehat{Ca}=0.04$ \cite{cerro-bubble-train},
$\widehat{Ca}=0.1$
\cite{cerro-space,wang-non-circular}, $\widehat{Ca}=0.033$ \cite{heil-threedim}. If the capillary
number is larger
than
the critical capillary number, i.e. $Ca>\widehat{Ca}$, then the bubble becomes axisymmetric with the
radius of the droplet dependant on the capillary number. One of the example for the bubble radii
dependence on the capillary number is in Fig. \ref{fig:heil:three:dim}.

@@

For simplicity, the periodic boundary conditions are applied. Thus, the bubble train is simulated.
To avoid the mutual influence of bubbles on each other the channel length was chosen as $15
H_{\mathrm{eff}}$. Periodic boundary conditions imply as well that the fluid is accelerated by a
body force in the framework of the LBM but not with the pressure difference. Therefore, we can not
address simulations with upward and downward flows \cite{cerro-bubble-train}. It is assumed that
the gravity effects are negligible and the flow is in the horizontal direction driven by the body
force. 
\begin{figure}[h]
\includegraphics*[bb=153 610 405 717,width=\textwidth]{Figures/benchmark_classical.eps} 
\caption{The classical benchmark layout is presented . The semiinfinite gas bubble
propagates through the liquid media. The inlet pressure is specified $P_{in}$, the outlet is the
free outflow. The bubble propagates with constant velocity $U_{\mathrm{bubble}}$. 
\label{fig:classical:benchmark}}
\end{figure}
\begin{figure}[h]
\includegraphics*[bb=152 470 410 713,width=\textwidth]{Figures/benchmark_lbm.eps}
\caption{The lattice Boltzmann benchmark layout is used in present calculcations. The dimensions of
the domain are chosen as $H_{\mathrm{eff}}\times H_{\mathrm{eff}}\times 15 H_{\mathrm{eff}}$. The
length of the bubble was chosen as $5 H_{\mathrm{eff}}$ for the film thickness to stabilize.
Periodic boundary conditions are applied in the streamwise direction. The flow is driven by body
force. \label{fig:lbm:benchmark}}
\end{figure}

@@

To further reduce the computational overhead only quarter of the channel is simulated, see Appendix
\ref{append:sym}. The next section briefly explains the binary liquid lattice Boltzmann model. Then
the steady-state approach, the dependency of the radii on the moderate capillary number values
$0.1\leq Ca \leq 1.0$, the velocity pattern, the relative bubble velocity and the aspect ratio are
examined, see Section \ref{sec:results}.

@@

The set of equations (\ref{set:equilibrium:binary}) restores the macroscopic
fluid equations as:
\begin{equation}
\label{full:navier:stokes}
\begin{aligned}
&\partial_t \rho+ \partial_{\alpha} \rho u_{\alpha}=0\\
&\rho\left(\partial_t+u_{\beta}\partial_{\beta}\right) u_{\alpha}=
-\partial_{\beta}P_{\alpha \beta} +
\nu\partial_{\beta}\left(\rho\partial_{\alpha}u_{\beta}+\rho\partial_{\beta} u_{\alpha}\right)\\
&\partial_t \phi + \partial_{\alpha} \phi u_{\alpha}=M \partial^2_{\beta\beta} \mu,
\end{aligned}
\label{binary:fluid:system}
\end{equation}
where $\nu=c_s^2 (\tau-1/2)$ is the dynamic viscosity,
$M=\Gamma(\tau_{\phi}-1/2)$ is the mobility parameter, and $\tau=\frac{1}{\omega}$ and $\tau_{\phi}$
are the relaxation parameters of density and phase fields.

@@

To further examine the steady-state the velocity in the streamwise direction is plotted. The values
of the velocity are taken on the contour where phase values are taken as $\phi=0$. This corresponds
 to the bubble interface. One can see the contour plot and the corressponing streamwise velocity
component in Fig. \ref{fig:velocity:contour}. The crossection is a plane $x=0$, where $x$ points
towards the reader, $z$ is the streamwise direction. As far as fluxes inside the bubble has
clockwise and counterclockwise directions, see Section \ref{sec:velocity}, one needs to thoroughly
examine only the points corresponding to the front and the rear bubble tips. If one finds it equal
to each other, then it's exactly the velocity with which the bubble propagates in the microchannel. 
\begin{figure}[h!]
\includegraphics[width=\textwidth]{Figures/velocity_interface_contour.eps}\\
\includegraphics[width=\textwidth]{Figures/velocity_interface_values.eps}\\
\caption{The bubble shape and the corresponding streamwise velocities taken on the bubble
interface after $240000$ iterations. Directions $y$ and $z$ are scaled to the half of the channel
$H_{\mathrm{eff}}/2$. One can see that the bubble front and rear tips are propagating with the
same velocities, which is actually the bubble velocity. For more details on velocity patterns
see Section \ref{sec:velocity}. \label{fig:velocity:contour}}
\end{figure}

@@

\section{Velocity pattern}
\label{sec:velocity}
The velocity pattern is known to change its behavior depending on the capillary number. For
relatively low capillary number $Ca<0.6$ there exists a vortex in front of the bubble in the moving
frame moving with the bubble \cite{heil-threedim}. For larger ones it is indicated that there is no
vortex in front of the bubble. The transition capillary number is indicated by
\citet{heil-threedim} and it equals for the rectangular channel $Ca=0.691$. However, the
distribution of vorticity strongly depends on Reynolds number \cite{heil-bretherton}, as well the
pressure distribution. In the case of the bubble train it is indicated by \citet{kreutzer-taylor}
that the pressure is largely influenced by bubble frequency and slug distance. The present
simulations are conducted for bubble trains and certain differences in terms of changing streamline
patterns are expected. We examined velocity patterns to identify the moment of the streamlines
pattern change. We chose two
representative capillary numbers as $Ca=0.47$ and $Ca=0.63$. 
One can
see a clear transition between associated patterns. The transition capillary number
\cite{heil-threedim} is slightly higher than the simulation results. In present simulations the
transition occurs at $Ca=0.5-0.6$.
\begin{figure}[h]
\includegraphics[angle=90,width=\textwidth]{Figures/stream_ca47.eps}\\
\includegraphics[angle=90,width=\textwidth]{Figures/stream_ca63.eps}\\
\caption{Velocity vector maps for $Ca=0.47$ and $Ca=0.63$. One can see that there are no vortexes
created
before the bubble for $Ca=0.63$. The transition happens between $Ca=0.47$ and $Ca=0.63$, which
is a different value from $Ca=0.69$ \cite{heil-threedim}. \label{fig:streamlines:pattern}}
\end{figure}

\section{Variation bubble length}
The work of \citet{wang-non-circular} shows the variation of the bubble
thickness through the length of
bubble. The bubble shape in terms of axial and diagonal radii is reconstructed for a number of
capillary numbers, Fig. \ref{fig:bubble:variation:capillaries}. One can see that for smaller
capillary numbers the diagonal radius exhibits a small jump near the rear bubble tip. It coincides
with results of \citet{wang-non-circular}. However, present simulations show smoother behavior of
the jump.  
\begin{figure}[h]
\includegraphics[width=\textwidth]{Figures/bubble_rad_axis.eps}\\
\includegraphics[width=\textwidth]{Figures/bubble_rad_diag.eps}\\
\caption{Radii variations along the bubble for different capillary numbers in the plane $x=0$.
One can see that the diagonal radius shows the jump near the bubble rear tip. It coincides with the
VOF simulations of \citet{wang-non-circular}. \label{fig:bubble:variation:capillaries}}
\end{figure}

Another interesting phenomenon for the bubble shape was indicated by \citet{heil-threedim}. In
their simulations the non-symmetric shape was observed even for larger capillary number $Ca>4$ near
the bubble front tip. We performed the same simulations.

\begin{figure}
\includegraphics[width=0.97\textwidth]{Figures/variationoverlength.eps}
\caption{The variation of the bubble length for capillary number $Ca=10$. The x-axis scaled to
half-width of the microchannel height.  \label{fig:thickness:variation:ca:ten}}
\end{figure}

Concerning results indicated in Fig.
\ref{fig:thickness:variation:ca:ten} simulations are unstable that means one needs to work with
small surface tension ({\color{red} See results of Pagonabarraga}). One can see that the difference
between diagonal radius and
usual radius is of order of $4\%$. That means it's 2 percent of the channel height. If the
simulation is conducted by the grid $100\mathrm{x}100$ in the cross section that means that we will
see the difference in two grid nodes. Therefore the grid is $102\mathrm{x}102\mathrm{1500}$. It is
known from simulations that for the grid $102\mathrm{x}102\mathrm{x}1500$ the largest force gradient
is $2\mathrm{x}10^{-5}$ with free energy binary liquid parameters $A=0.04$,$k=0.04$ and $Ca\approx
1.0$. Therefore, we want to keep the same body force but to obtain larger capillary number. That
can be done through the scaling indicated in Appendix \ref{append:scaling}:
\begin{equation}
\frac{\mathrm{d}P}{\mathrm{d}x}\propto \gamma Ca .
\end{equation}
Therefore the same body force can stand for capillary number ten times larger if the surface
tension is ten times smaller:
\begin{equation}
\gamma=\sqrt{\frac{8 k A}{9}}.
\end{equation}
The following can be satisfied if $A=0.004$ and $k=0.004$. Decreasing surface tension requires
increasing the mobility of the interface to be able to adjust to the interface
\cite{pagonabarraga-parameters}. We therefore performed the numerical simulation with those
parameters and $\Gamma=20$ for the interface to be stable. 
{\color{red} However, many authors state that increasing mobility kills
the
proper results of simulations}.
{\color{red} Need to wait results from bugaboo.westgrid.ca under LargeCap simulation}

\section{Capillary number}

While changing the applied force in a manner indicated in Appendix
\ref{append:scaling} the velocity of the bubble was increased to obtain larger capillary numbers,
see Fig. \ref{fig:underresolved:capillaries}. One can see that the deviation between the diagonal
and axial radiuses are quite significant and they coincide only when the interface is resolved
($Ca\approx 1.0$). In calculations with better grids, i.e. quarter channels simulations on CPU, the
axial radius coincides with the diagonal in the $0.5\%$ range. Moreover, the capillary number curve
resembles already published date of \citet{heil-threedim} in a better way, see Fig.
\ref{fig:capillary:comparison}.

We performed a number of simulations with quarter CPU code and GPU train simulations and compared
them with data obtained by \cite{heil-threedim}. After performing the simulations we determine the
$R_{axis}$, $R_{diag}$ versus calculated
capillary number
\begin{equation}
Ca=\frac{\mu_{liq} U_{bubble}}{\gamma}.
\end{equation}
To perform simulations we
started with the following parameters as
$k=0.04$,$A=0.04$,$\Gamma=1.0$. We performed different grids calculations with GPU and CPU codes.
The GPU code (see section \ref{section:bubble:train} about bubble train simulations) simulates a
quarter of the bubble and operates on the grid $100\mathrm{x}100\mathrm{x}1500$. The CPU refined
grid operates on whole $160\mathrm{x}160\mathrm{x}1500$. In this case we see short bubbles as the
length to height ratio is $\frac{500}{160}=3.125$. GPU  For GPU grids due to memory constrains we
chose the grid as $N_x\mathrm{x}N_y\mathrm{x}N_z=52\mathrm{x}52\mathrm{x}750$. One can see the
comparison presented in Fig. \ref{fig:capillary:comparison}.
\begin{figure}
\includegraphics[width=0.97\textwidth]{Figures/capillaries_comparison.eps}
\caption{The comparison between different runs of codes for the axial and the diagonal radiuses
versus capillary numbers. \label{fig:capillary:comparison}}. One can see that the code mimics
behavior of the earlier published results.
\end{figure}

\section{Train simulation}
\label{section:bubble:train}
As far as the requirements for the grid are ``heavy'' we performed a bunch of simulations to
examine how the distance between bubbles influences on the thickness. For this purpose the
simulation was conducted on the grid 82x82x750. We examined different initialized distances between
bubbles. The initial lengths of bubbles were chosen as $300$, $350$, $400$, $450$, $500$, $550$
lattice units. Therefore, the corresponding initial distances between bubbles are
$450$,$400$,$350$,$300$,$250$ lattice units. On the steady state regime the distances become $386$,
$317$, $248$,$176$,$99$,$23$ lattice units. The corresponding capillary numbers calculated based on
the center bubble velocity are as follows $0.905$, $0.986$,$1.08$,$1.19$, $1.33$, $1.49$. The
calculated radiuses (diagonal radius equal to the axes radius) are as follows
$0.783$,$0.778$,$0.772$,$0.764$,$0.753$,$0.747$. Thus, one can see the great impact of the bubbles
on each other velocity for the given body force. However, in terms of the film thicknesses those
numbers look adequate. One can see the film thicknesses dependency on the capillary number for
simulations obtained by \citet{heil-threedim} and bubble train simulations data in Fig.
\ref{fig:capillaries:train} and in Fig. \ref{fig:capillary:comparison}. One can see that the
results are consistent and in a good agreement.
Therefore, though the bubble train mutual motion influences on the film thickness by increasing the
associated capillary number,  however, the film thickness is in accordance with the capillary
number correlation. Thus, the decrease of the numerical domain is quantified and can be used to
obtain the data. However, one should be cautious about the initialization, because the Poiseuille
flow initialization is deviated even further due to mutual motion of bubbles.
\begin{figure}
\includegraphics[width=0.97\textwidth]{Figures/capillaries_comparison_train.eps}
\caption{The comparison for the film thickness between \citet{heil-threedim} data and bubble train
simulations. Though, one can see that the decrease of the distance between bubbles causes the
increase in the associated capillary number, however, the correlation dependance between film
thickness and capillary number stays the same. This can help for simulations to reduce the
numerical domain. \label{fig:capillaries:train}}
\end{figure}

\section{Conclusion}
The present work conducts the binary-liquid simulations with the lattice Boltzman method. The
method is a continuous interface method. Along with similarity of results with already published
results of \citet{heil-threedim} and \citet{wang-non-circular}, the differencies were found,
especially in the velocity pattern. Results evedince that the lattice Boltzmann binary liquids
simulations can be used for simulation of gas bubbles in the microchannel.

@@

>>>>

In comparison with two-dimensional geometries, the experimental works
\cite{shikazono-square,cerro-bubble-train} supported by numerical simulations \cite{heil-threedim,
wang-non-circular} show interesting phenomena in three-dimensional geometries. It was found
\cite{heil-threedim,wong-films} that for rectangular
or square shaped capillaries there is a transition for a
certain capillary number, where the bubble crosssection changes from the non axisymmetric to axis
symmetric case. Non axisymmetric case is attributed to the case where the axial radius is
different from the
diagonal radius and the bubble has the non-circular shape in the channel crossection. In this case
the bubble shape mimics the shape of the square and looks like a rounded square. The dependence
of the diagonal and axial radii on the capillary number is shown in Fig. 
\ref{fig:heil:three:dim}.
\begin{figure}[ht]
\includegraphics[width=\textwidth]{Figures/capillary_width_heil.eps}
\caption{\citet{heil-threedim} results for the variation of the deposited liquid for the range of
capillary number. One can see the asymmetry between diagonal and axis diameter for the capillary
number $Ca\leq\hat{Ca}=0.04$. Courtesy of \citet{heil-threedim}. \label{fig:heil:three:dim}}
\end{figure}
The transition number between non-axisymmetric case and symmetric is indicated in a number of
works as $\widehat{Ca}=0.04$ \cite{cerro-bubble-train},
$\widehat{Ca}=0.1$
\cite{cerro-space,wang-non-circular}, $\widehat{Ca}=0.033$ \cite{heil-threedim}. If the capillary
number is larger
than
the critical capillary number, i.e. $Ca>\widehat{Ca}$, then the bubble becomes axisymmetric with the
radius of the droplet dependant on the capillary number. One of the example for the bubble radii
dependence on the capillary number is in Fig. \ref{fig:heil:three:dim}.

@@

For simplicity, the periodic boundary conditions are applied. Thus, the bubble train is simulated.
To avoid the mutual influence of bubbles on each other the channel length was chosen as $15
H_{\mathrm{eff}}$. Periodic boundary conditions imply as well that the fluid is accelerated by a
body force in the framework of the LBM but not with the pressure difference. Therefore, we can not
address simulations with upward and downward flows \cite{cerro-bubble-train}. It is assumed that
the gravity effects are negligible and the flow is in the horizontal direction driven by the body
force. 
\begin{figure}[ht]
\includegraphics*[bb=153 610 405 717,width=\textwidth]{Figures/benchmark_classical.eps} 
\caption{The classical benchmark layout is presented . The semiinfinite gas bubble
propagates through the liquid media. The inlet pressure is specified $P_{in}$, the outlet is the
free outflow. The bubble propagates with constant velocity $U_{\mathrm{bubble}}$. 
\label{fig:classical:benchmark}}
\end{figure}
\begin{figure}[ht]
\includegraphics*[bb=152 470 410 713,width=\textwidth]{Figures/benchmark_lbm.eps}
\caption{The lattice Boltzmann benchmark layout is used in present calculcations. The dimensions of
the domain are chosen as $H_{\mathrm{eff}}\times H_{\mathrm{eff}}\times 15 H_{\mathrm{eff}}$. The
length of the bubble was chosen as $5 H_{\mathrm{eff}}$ for the film thickness to stabilize.
Periodic boundary conditions are applied in the streamwise direction. The flow is driven by body
force. \label{fig:lbm:benchmark}}
\end{figure}

@@

To further reduce the computational overhead only quarter of the channel is simulated, see Appendix
\ref{append:sym}. The next section briefly explains the binary liquid lattice Boltzmann model. Then
the steady-state approach, the dependency of the radii on the moderate capillary number values
$0.1\leq Ca \leq 1.0$, the velocity pattern, and the relative bubble velocity, see Section
\ref{sec:results}.

@@

The set of equations (\ref{set:equilibrium:binary}) restores the macroscopic
fluid equations as:
\begin{equation}
\label{full:navier:stokes}
\begin{aligned}
&\partial_t \rho+ \partial_{\alpha} \rho u_{\alpha}=0\\
&\rho\left(\partial_t+u_{\beta}\partial_{\beta}\right) u_{\alpha}=
-\partial_{\beta}P_{\alpha \beta} +
\nu\partial_{\beta}\left(\rho\partial_{\alpha}u_{\beta}+\rho\partial_{\beta} u_{\alpha}\right)\\
&\partial_t \phi + \partial_{\alpha} \phi u_{\alpha}=M \partial^2_{\beta\beta} \mu,
\end{aligned}
\end{equation}
where $\nu=c_s^2 (\tau-1/2)$ is the dynamic viscosity,
$M=\Gamma(\tau_{\phi}-1/2)$ is the mobility parameter, and $\tau=\frac{1}{\omega}$ and $\tau_{\phi}$
are the relaxation parameters of density and phase fields.

@@

To further examine the steady-state the velocity in the streamwise direction is plotted. The values
of the velocity are taken on the contour where phase values are taken as $\phi=0$. This corresponds
 to the bubble interface. One can see the contour plot and the corressponing streamwise velocity
component in Fig. \ref{fig:velocity:contour}. The crossection is a plane $x=0$, where $x$ points
towards the reader, $z$ is the streamwise direction. As far as fluxes inside the bubble has
clockwise and counterclockwise directions, see Section \ref{sec:velocity}, one needs to thoroughly
examine only the points corresponding to the front and the rear bubble tips. If one finds it equal
to each other, then it's exactly the velocity with which the bubble propagates in the microchannel.
For physical analogy, one can imagine the rotating ball propagating in the streamwise direction. 
\begin{figure}[ht]
\includegraphics[width=\textwidth]{Figures/velocity_interface_contour.eps}\\
\includegraphics[width=\textwidth]{Figures/velocity_interface_values.eps}\\
\caption{The bubble shape and the corresponding streamwise velocities taken on the bubble
interface after $240000$ iterations. Directions $y$ and $z$ are scaled to the half of the channel
$H_{\mathrm{eff}}/2$. One can see that the bubble front and rear tips are propagating with the
same velocities, which is actually the bubble velocity. For more details on velocity patterns
see Section \ref{sec:velocity}. \label{fig:velocity:contour}}
\end{figure}

@@

\section{Velocity pattern}
\label{sec:velocity}
The velocity pattern is known to change its behavior depending on the capillary number. For
relatively low capillary number $Ca<0.6$ there exists a vortex in front of the bubble in the moving
frame moving with the bubble \cite{heil-threedim}. For larger ones it is indicated that there is no
vortex in front of the bubble. The transition capillary number is indicated by
\citet{heil-threedim} and it equals for the rectangular channel $Ca=0.691$. However, the
distribution of vorticity strongly depends on Reynolds number \cite{heil-bretherton}, as well the
pressure distribution. In the case of the bubble train it is indicated by \citet{kreutzer-taylor}
that the pressure is largely influenced by bubble frequency and slug distance. The present
simulations are conducted for bubble trains and certain differences in terms of changing streamline
patterns are expected. We examined velocity patterns to identify the moment of the streamlines
pattern change. We chose two
representative capillary numbers as $Ca=0.47$ and $Ca=0.63$. 
One can
see a clear transition between associated patterns. The transition capillary number
\cite{heil-threedim} is slightly higher than the simulation results. In present simulations the
transition occurs at $Ca=0.5-0.6$.
\begin{figure}[ht]
\includegraphics[angle=90,width=\textwidth]{Figures/stream_ca47.eps}\\
\includegraphics[angle=90,width=\textwidth]{Figures/stream_ca63.eps}\\
\caption{Velocity vector maps for $Ca=0.47$ and $Ca=0.63$. One can see that there are no vortexes
created
before the bubble for $Ca=0.63$. The transition happens between $Ca=0.47$ and $Ca=0.63$, which
is a different value from $Ca=0.69$ \cite{heil-threedim}. \label{fig:streamlines:pattern}}
\end{figure}

\section{Variation bubble length}
The work of \citet{wang-non-circular} shows the variation of the bubble
thickness through the length of
bubble. The bubble shape in terms of axial and diagonal radii is reconstructed for a number of
capillary numbers, Fig. \ref{fig:bubble:variation:capillaries}. One can see that for smaller
capillary numbers the diagonal radius exhibits a small jump near the rear bubble tip. It coincides
with results of \citet{wang-non-circular}. However, present simulations show smoother behavior of
the jump.  
\begin{figure}[ht]
\includegraphics[width=0.5\textwidth]{Figures/bubble_rad_axis.eps}
\includegraphics[width=0.5\textwidth]{Figures/bubble_rad_diag.eps}\\
\caption{Radii variations along the bubble for different capillary numbers in the plane $x=0$.
One can see that the diagonal radius shows the jump near the bubble rear tip. It coincides with the
VOF simulations of \citet{wang-non-circular}.Coordinate $z$ increases in the streamwise direction.
\label{fig:bubble:variation:capillaries}}
\end{figure}

Another interesting phenomenon for the bubble shape was indicated by \citet{heil-threedim}. In
their simulations the non-symmetric shape was observed even for larger capillary number $Ca>4$ near
the bubble front tip which after certain distance decay to have radii equal to each other. The
bubble shape was calculated for the large capillary number $Ca=6.43$, see Fig.
\ref{fig:bubble:ca:large}. However, even the diagonal radius shows the jump near the front tip of
the bubble and it resembles the shape given by \citet{heil-threedim}, but the difference is around
$1$ percent and can be attributed to the linear interpolation error used for the interface
calculation. 
\begin{figure}[ht]
\includegraphics[width=\textwidth]{Figures/bubble_ca_large.eps}
\caption{The bubble shape for the capillary number $Ca=6.43$ is presented.$z$ increases in the
streamwise direction. One can see that near the front tip (large $z$) the diagonal radius exhibits
a jump which is relaxed after to the axis radius. Even this resembles
the shape of the curves obtained by \citet{heil-threedim}, but the
difference between radii is around $1$ percent and can be attributed to
the linear interpolation used in the calculation of the interface.\label{fig:bubble:ca:large}}
\end{figure}

\section{Capillary number}
The purpose of this section is to compare radii dependence on the capillary number. The film
thicknesses were extracted from the middle of the bubble. The data presented in Fig.
\ref{fig:capillary:comparison} is simulated with different techniques including different
initialization, different grids as $52 \times 52 \times 1500$, $82 \times 82 \times 1500$ and $82
\times 82 \time 2400$. All the results are consistent. Therefore for the capillary number $Ca\geq
0.05$ the grid independence is achieved. The results are compared with results of
\citet{heil-threedim} and of \citet{wang-non-circular}. The data was extracted from the
correspondent references with the help of the program ``Engauge Digitizer''. The lattice Boltzmann
can catch the transition between non-symmetric to axisymmetric shape of the bubble.

Present simulations show that the corresponding critical capillary number is close to
$\widehat{Ca}=0.1$ which is close to work of \citet{wang-non-circular}. 
In comparison with
the continuous method VOF \cite{wang-non-circular} present simulations show less noisy trend and
more accurate transition.

For moderate capillary number $Ca\geq 0.2$ radii show less value than the literature results. That
might be corresponding to the bubble train simulation or the influence of the binary-liquid model. 
\begin{figure}[ht]
\includegraphics[width=0.5\textwidth]{Figures/capillaries_comparison_heil.eps}
\includegraphics[width=0.5\textwidth]{Figures/capillaries_comparison_wang.eps}\\
\caption{The comparison for the axial and the diagonal radii
versus capillary numbers between present simulations and the results of \citet{heil-threedim}
(top), of \citet{wang-non-circular} (bottom). One can see that
the code mimics
behavior of the earlier published results.  \label{fig:capillary:comparison}} 
\end{figure}

\section{Relative velocity}
One of the differencies between the flow in tubes and square cappillaries is that the flow has
different patterns. In square microchannels the main mass flow goes through corners
\cite{heil-threedim,wang-non-circular}. This affects the mass flow significantly. One of the
characteristics of the mass flow is the normalized relative velocity dependence on the capillary
number. Bubble moves always faster than surrounding liquid. The following quantity is defined as
relative velocity \cite{cerro-bubble-train}:
\begin{equation}
W=\frac{U_{\mathrm{bubble}}-U_{ls}}{U_{\mathrm{bubble}}},
\end{equation}
where $U_{\mathrm{bubble}}$ is the bubble interface velocity, $U_{ls}$ is the liquid superficial
velocity. As far as the steady state is achieved one can take the liquid superficial velocity
exactly at the distance in the middle between bubbles. Then, the liquid superficial velocity is
defined as:
\begin{equation}
U_{ls}=\frac{\int_{A}{U_{liq}dA}}{A}.
\end{equation}
The relative velocity was calculated and compared with the results of \citet{wang-non-circular}.

% \section{Train simulation}
% \label{section:bubble:train}
% As far as the requirements for the grid are ``heavy'' we performed a bunch of simulations to
% examine how the distance between bubbles influences on the thickness. For this purpose the
% simulation was conducted on the grid 82x82x750. We examined different initialized distances
%between
% bubbles. The initial lengths of bubbles were chosen as $300$, $350$, $400$, $450$, $500$, $550$
% lattice units. Therefore, the corresponding initial distances between bubbles are
% $450$,$400$,$350$,$300$,$250$ lattice units. On the steady state regime the distances become
%$386$,
% $317$, $248$,$176$,$99$,$23$ lattice units. The corresponding capillary numbers calculated based
%on
% the center bubble velocity are as follows $0.905$, $0.986$,$1.08$,$1.19$, $1.33$, $1.49$. The
% calculated radiuses (diagonal radius equal to the axes radius) are as follows
% $0.783$,$0.778$,$0.772$,$0.764$,$0.753$,$0.747$. Thus, one can see the great impact of the bubbles
% on each other velocity for the given body force. However, in terms of the film thicknesses those
% numbers look adequate. One can see the film thicknesses dependency on the capillary number for
% simulations obtained by \citet{heil-threedim} and bubble train simulations data in Fig.
% \ref{fig:capillaries:train} and in Fig. \ref{fig:capillary:comparison}. One can see that the
% results are consistent and in a good agreement.
% Therefore, though the bubble train mutual motion influences on the film thickness by increasing
%the
% associated capillary number,  however, the film thickness is in accordance with the capillary
% number correlation. Thus, the decrease of the numerical domain is quantified and can be used to
% obtain the data. However, one should be cautious about the initialization, because the Poiseuille
% flow initialization is deviated even further due to mutual motion of bubbles.
% \begin{figure}
% \includegraphics[width=0.97\textwidth]{Figures/capillaries_comparison_train.eps}
% \caption{The comparison for the film thickness between \citet{heil-threedim} data and bubble train
% simulations. Though, one can see that the decrease of the distance between bubbles causes the
% increase in the associated capillary number, however, the correlation dependance between film
% thickness and capillary number stays the same. This can help for simulations to reduce the
% numerical domain. \label{fig:capillaries:train}}
% \end{figure}

\section{Conclusion}
The present work conducts the binary-liquid three-dimensional simulations with the lattice Boltzman
method. The
method is a continuous interface method. The results are similar with the literature  Along with
similarity of results with already published
results of \citet{heil-threedim} and \citet{wang-non-circular}, the differencies were found,
especially in the velocity pattern. Results evedince that the lattice Boltzmann binary liquids
simulations can be used for simulation of gas bubbles in the microchannel.

@@

can be done through the scaling indicated in Appendix \ref{append:scaling}:
\begin{equation}
\frac{\mathrm{d}P}{\mathrm{d}x}\propto \gamma Ca .
\end{equation}
Therefore the same body force can stand for capillary number ten times larger if the surface
tension is ten times smaller:
\begin{equation}
\gamma=\sqrt{\frac{8 k A}{9}}.
\end{equation}
The following can be satisfied if $A=0.004$ and $k=0.004$. Decreasing surface tension requires
increasing the mobility of the interface to be able to adjust to the interface
\cite{pagonabarraga-parameters}. We therefore performed the numerical simulation with those
parameters and $\Gamma=20$ for the interface to be stable. 
{\color{red} However, many authors state that increasing mobility kills
the
proper results of simulations}.
{\color{red} Need to wait results from bugaboo.westgrid.ca under LargeCap simulation}

To perform simulations we
started with the following parameters as
%k=0.04$,$A=0.04$,$\Gamma=1.0$. We performed different grids calculations with GPU and CPU codes.
%$The GPU code (see section \ref{section:bubble:train} about bubble train simulations) simulates a
%quarter of the bubble and operates on the grid $100\mathrm{x}100\mathrm{x}1500$. The CPU refined
%grid operates on whole $160\mathrm{x}160\mathrm{x}1500$. In this case we see short bubbles as the
%length to height ratio is $\frac{500}{160}=3.125$. GPU  For GPU grids due to memory constrains we
%chose the grid as $N_x\mathrm{x}N_y\mathrm{x}N_z=52\mathrm{x}52\mathrm{x}750$. One can see the
%comparison presented in Fig. \ref{fig:capillary:comparison}.
